Parsing Postfix Log Files

### 2009-04-23 doc/logparser-conclusion.tex
Changes made based on Carl's comments.

The flow of control in this architecture is quite different from most other
architectures.  Most traditional parsers have a current state: each state
has a fixed set of acceptable next states, processing is determined by the
state transition that takes place, and unacceptable state transitions cause
parsing to fail.  This architecture is different: the rule that recognises
the input dictates the action that will be invoked.  Rule conditions
(\sectionref{rule conditions in architecture}) enable stateful parsing,
where the list of rules used to recognise an input is constrained by the
parser's current state, but the recognising rule still dictates the action
that is invoked and, whether directly or indirectly, the next state.

@@

A parser's ability to correctly parse its inputs is extremely important;
\parsernames{} coverage of \numberOFlogFILES{} log files is discussed in
\sectionref{parsing coverage}.  Both its success at recognising individual
log lines  and its correctness in reconstructing each mail's journey
through Postfix are described in detail, including the results of manually
verifying that a randomly selected portion of a log file was correctly
parsed.  Experience implementing \parsername{} shows that full input
coverage is relatively easy to achieve with this architecture, and that
with enough time and effort a full understanding of the input is possible.
Postfix log files would require substantial time and effort to correctly
parse regardless of the architecture used; this architecture enables an
iterative approach to be used~\cite{stepwise-refinement}, as is practiced
in many other software engineering disciplines.

>>>>

The flow of control in this architecture is quite different from other
architectures, e.g.\ those used for compiling a programming language.
Typically, those parsers have a current state: each state has a fixed set
of acceptable next states, processing is determined by the state transition
that takes place, and unacceptable state transitions cause parsing to fail.
This architecture is different: the rule that recognises the input dictates
the action that will be invoked.  Rule conditions (\sectionref{rule
conditions in architecture}) enable stateful parsing, where the list of
rules used to recognise an input is constrained by the parser's current
state, but the recognising rule still dictates the action that is invoked
and, whether directly or indirectly, the next state.

@@

A parser's ability to correctly parse its inputs is extremely important;
\parsernames{} coverage of \numberOFlogFILES{} log files, each containing
one day's log lines, is discussed in \sectionref{parsing coverage}.  Both
its success at recognising individual log lines  and its correctness in
reconstructing each mail's journey through Postfix are described in detail,
including the results of manually verifying that a randomly selected
portion of a log file was correctly parsed.  Experience implementing
\parsername{} shows that full input coverage is relatively easy to achieve
with this architecture, and that with enough time and effort a full
understanding of the input is possible.  Postfix log files would require
substantial time and effort to correctly parse regardless of the
architecture used; this architecture enables an iterative approach to be
used~\cite{stepwise-refinement}, as is practiced in many other software
engineering disciplines.

### 2009-04-23 doc/logparser-results.tex
Changes made based on Carl's comments.

Efficiency is an obvious concern when the parser routinely needs to parse
large log files.  The mail server that generated the \numberOFlogFILES{}
log files is a production mail server handling mail for a university
department; the benefit of using this mail server is that its log files
exhibit the idiosyncrasies and peculiarities that a mail server in the wild
must deal with, but the downside is that significantly altering the
configuration to accommodate this project is not an option.
\Graphref{Mails received per day} and \tableref{Number of mails received
per day: statistics} show that, as expected, it accepts far more mails on
weekdays than at weekends.  Note that these figures only count mails
received by \acronym{SMTP}, and the mail loops noticeable in later graphs
do not affect the figures.

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling \input{build/include-total-size-of-93-log-files},
\numberOFlogLINEShuman{} log lines) were each parsed 10 times, and the mean
parsing time calculated.  The computer used for test runs was a Dell
Optiplex 745, shown in \tableref{Details of the computer used to generate
statistics}; it was dedicated to the task of gathering statistics from test
runs, and did not run any other programs while test runs were in progress.
Saving results to the database was disabled for the test runs, because that
dominates the run time of the parser, and the tests are aimed at measuring
the speed of \parsername{} rather than the speed of the database and the
disks it is stored on.  Parsing all \numberOFlogFILES{} log files in one
run without saving results to the database took
\input{build/include-full-run-duration}, with mean throughput of
\input{build/include-full-run-throughput}.  In contrast, when saving
results to the database, parsing all \numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput} --- parsing takes up only
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database.

@@

The \numberOFlogFILES{} log files used when generating statistics have
fairly consistent sizes and contents, except for two groups of log files:
22 \& 62--68.  Median log file size is
\input{build/include-median-file-size}, containing
\input{build/include-median-number-of-log-lines} log lines, with a median
of \input{build/include-median-mails-accepted-over-smtp} mails accepted
each weekday; large scale mail servers would generate much larger log
files.  \Graphref{parsing time vs log file size vs number of log lines
graph} shows the parsing time in seconds, log file size in MB, and number
of log lines in tens of thousands, for each of the \numberOFlogFILES{} log
files.  \Graphref{parsing time vs log file size vs number of log lines
factor} plots both the ratio of log file size vs parsing time, and the
ratio of number of log lines vs parsing time (higher is more efficient in
both cases); \tableref{parsing time vs log file size vs number of log lines
factor table} shows the same ratios for different groups of log files.  The
ratios are quite tightly banded, except they increase (i.e.\ improve) for
log files 22 \& 62--68, despite their larger than usual size.  Both groups
of log files are much larger than usual because of mail loops caused by
users who set up mail forwarding incorrectly, resulting in a very different
distribution of log lines: normally most log lines are generated by mail
delivery attempts from other hosts, but when the mail loops occurred most
of the log lines resulted from failed delivery of mail generated on the
server itself.  The Postfix components that generated most of the log lines
during the mail loops have fewer associated rules than the Postfix
components whose log lines normally make up the bulk of each log file, so
the mean number of rules tried per log line is lower for these log files,
as is the mean parsing time per log line.  \Tableref{Number of rules per
Postfix component} shows the number of rules for each Postfix component;
\graphref{Mean number of rules used per log line} shows the drop in the
mean number of rules tried per log line for log files containing a mail
loop, and \graphref{Mean number of rules used per log line for each Postfix
component} shows the mean number of rules tried per log line for each
Postfix component for all log files.

@@

@@

@@

\Graphref{Parsing time of optimal and reverse orderings relative to
shuffled ordering graph} shows the parsing times of optimal and reverse
orderings relative to shuffled ordering; the mean relative parsing times
for different groupings of log files are given in \tableref{Parsing time of
optimal and reverse orderings relative to shuffled ordering table}.  This
optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
\Tableref{Parsing time of optimal and reverse orderings relative to
shuffled ordering table} shows that differences in rule ordering have less
effect on parsing time when parsing log files 22 \& 62--68, because of the
different distribution of log lines in those log files.  A careful
examination of \graphref{Parsing time of optimal and reverse orderings
relative to shuffled ordering graph} shows that, for the first log file
only, optimal and reverse orderings perform identically: this is because
the hits field of each rule is zero for the first log file, so optimal and
reverse orderings produce identical rule orderings.  For the first log
file, shuffled ordering is the most efficient of the three, but that is
accidental and cannot be relied upon.

@@

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important for this architecture because it is
anticipated that the typical parser will have a large ruleset.  The full
\parsername{} ruleset has \numberOFrules{} rules, whereas the minimum
ruleset required to parse the \numberOFlogFILES{} log files has
\numberOFrulesMINIMUM{} rules, \numberOFrulesMINIMUMpercentage{} of the
full ruleset.  The full ruleset is larger because \parsername{} is tested
with \numberOFlogFILESall{} log files; testing with more log files
increases the chance of finding bugs in the parser or new complications to
be overcome.  The \numberOFlogFILES{} log files were each parsed 10 times
using the minimum ruleset, and the mean parsing times compared to those
generated using the full ruleset: the percentage parsing time increase when
using the full ruleset instead of the minimal ruleset for optimal,
shuffled, and reverse orderings is shown in \graphref{Percentage parsing
time increase when using the maximum ruleset instead of the minimum
ruleset}, with mean and standard deviation in \tableref{Percentage parsing
time increase when using the maximum ruleset instead of the minimum ruleset
table}.

@@

The optimal ordering shows a mean increase of just
\input{build/include-full-ruleset-vs-minimum-ruleset-mean} in parsing time
for a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
These results show that both the architecture and \parsername{} scale
extremely well as the ruleset increases in size, and that optimally
ordering the rules is an important optimisation contributing to this
scalability.

@@

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the effect that not caching compiled regexes has on parser
performance, with mean and standard deviation in \tableref{Increase in
parsing time when not caching compiled regexes table}.  For typical log
files, the mean increase in parsing time when not caching compiled regexes
is \input{build/include-cached-regexes-vs-discarded-regexes-mean}.  Caching
compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement.  As seen
previously, log files 22 \& 62--68 do not suffer such a large increase in
parsing time when the optimisation is disabled; this is because, on
average, fewer rules are used so fewer regexes are compiled per log line
for those log files.  The increase in parsing time when parsing the first
log file is much greater than for the other log files; again, this is
because every rule's hits field starts at zero, so optimal ordering is less
efficient than usual, the mean number of rules used for each log line is
higher than usual, and more regexes will need to be compiled when
recognising each log line.

@@

The optimisations described in this chapter have optimised the process of
recognising log lines, but have not optimised actions at all.  Optimisation
efforts have concentrated on recognition of log lines for two reasons:

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still accounts for two thirds of the parser's
        execution time.  \Graphref{Percentage of parsing time spent
        recognising log lines graph} shows the percentage of parsing time
        spent recognising log lines for each of the \numberOFlogFILES{} log
        files, with mean and standard deviation shown in
        \tableref{Percentage of parsing time spent recognising log lines
        table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  Optimal ordering
        reduces parsing time by
        \input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
        relative to shuffled ordering; actions occupy
        \input{build/include-percentage-time-spent-processing-log-lines-mean}
        of parsing time, less if any optimisations are disabled, so
        optimising actions could not possibly provide as big a performance
        increase as optimally ordering rules.  Similarly, caching compiled
        regexes provides an
        \input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}
        reduction in parsing time; not invoking actions at all reduces the
        most optimised parsing time by less than half that.  If processing
        of recognised log lines was optimised to 1\% of its original
        parsing time, it would be only slightly more effective than optimal
        ordering, but it would be vastly harder to implement; optimising
        actions would not reduce parsing time enough to justify the amount
        of effort required.

>>>>

Efficiency is an obvious concern when the parser routinely needs to parse
large log files.  The mail server that generated the \numberOFlogFILES{}
log files is a production mail server handling mail for a university
department; the benefit of using this mail server is that its log files
exhibit the idiosyncrasies and peculiarities that a mail server in the wild
must deal with, but the downside is that significantly altering the
configuration to accommodate this project is not an option.
\Graphref{Mails received per day} and \tableref{Number of mails received
per day: statistics} show that, as expected, it accepts far more mails on
weekdays than at weekends.  Note that mail generated on the server does not
contribute to the figures: in particular, the vast amount of mail resulting
from the mail loops noticeable in later graphs is not reflected in these
figures.

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files were used, each containing one day's log lines, totalling
\input{build/include-total-size-of-93-log-files}, \numberOFlogLINEShuman{}
log lines.  Each log file was parsed 10 times, and the mean parsing time
calculated.  The computer used for test runs was a Dell Optiplex 745, shown
in \tableref{Details of the computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and did not
run any other programs while test runs were in progress.  Saving results to
the database was disabled for the test runs, because that dominates the run
time of the parser, and the tests are aimed at measuring the speed of
\parsername{} rather than the speed of the database and the disks it is
stored on.  Parsing all \numberOFlogFILES{} log files in one run without
saving results to the database took
\input{build/include-full-run-duration}, with mean throughput of
\input{build/include-full-run-throughput}.  In contrast, when saving
results to the database, parsing all \numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput} --- parsing takes up only
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database.

@@

The \numberOFlogFILES{} log files used when generating statistics have
fairly consistent sizes and contents, except for two groups of log files:
22 \& 62--68.  Median log file size is
\input{build/include-median-file-size}, containing
\input{build/include-median-number-of-log-lines} log lines, with a median
of \input{build/include-median-mails-accepted-over-smtp} mails accepted
each weekday; large scale mail servers would generate much larger log
files.  \Graphref{parsing time vs log file size vs number of log lines
graph} shows the parsing time in seconds, log file size in MB, and number
of log lines in tens of thousands, for each of the \numberOFlogFILES{} log
files.  Parsing time is plotted against number of log lines in
\graphref{Parsing time plotted against number of log lines}, and against
log file size in \graphref{Parsing time plotted against log file size};
the points on both graphs form almost perfectly straight lines.

\Graphref{parsing time vs log file size vs number of log lines
factor} plots both the ratio of log file size vs parsing time, and the
ratio of number of log lines vs parsing time (higher is more efficient in
both cases); \tableref{parsing time vs log file size vs number of log lines
factor table} shows the same ratios for different groups of log files.  The
ratios are quite tightly banded, except they increase (i.e.\ improve) for
log files 22 \& 62--68, despite their larger than usual size.  Both groups
of log files are much larger than usual because of mail loops caused by
users who set up mail forwarding incorrectly, resulting in a very different
distribution of log lines: normally most log lines are generated by mail
delivery attempts from other hosts, but when the mail loops occurred most
of the log lines resulted from failed delivery of mail generated on the
server itself.  The Postfix components that generated most of the log lines
during the mail loops have fewer associated rules than the Postfix
components whose log lines normally make up the bulk of each log file, so
the mean number of rules tried per log line is lower for these log files,
as is the mean parsing time per log line.  \Tableref{Number of rules per
Postfix component} shows the number of rules for each Postfix component;
\graphref{Mean number of rules used per log line} shows the drop in the
mean number of rules tried per log line for log files containing a mail
loop, and \graphref{Mean number of rules used per log line for each Postfix
component} shows the mean number of rules tried per log line for each
Postfix component for all log files.

The three plots on \graphref{parsing time vs log file size vs number of log
lines graph} are quite close together, and the graph can be difficult to
read.  Log files 62--68 show a large gap between log file size and number
of log lines, whereas those plots appear to have a small gap or to overlap
for all the other log files, suggesting that the ratio of log file size to
number of log lines is lower for log files 62--68.  This appearance is
misleading: \graphref{Mean log line size in bytes} shows the mean log line
size for all \numberOFlogFILES{} log files, and it reduces for log files 22
\& 62--68, i.e.\ there are more log lines per MB in those log files.

@@

\showgraph{build/graph-parsing-time-vs-file-size}{Parsing time plotted
against log file size}{Parsing time plotted against log file size}

\showgraph{build/graph-parsing-time-vs-number-of-log-lines}{Parsing time
plotted against number of log lines}{Parsing time plotted against number of
log lines}

@@

\showgraph{build/graph-mean-bytes-per-log-line}{Mean log line size in
bytes}{Mean log line size in bytes}

@@

The parsing time of the three orderings are plotted against log file size
in \graphref{Parsing time plotted against log file size for optimal,
shuffled, and reverse orderings}; \graphref{Parsing time of optimal and
reverse orderings relative to shuffled ordering graph} shows the parsing
times of optimal and reverse orderings relative to shuffled ordering, and
mean relative parsing times for different groupings of log files are given
in \tableref{Parsing time of optimal and reverse orderings relative to
shuffled ordering table}.  This optimisation provides a mean reduction in
parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
\Tableref{Parsing time of optimal and reverse orderings relative to
shuffled ordering table} shows that differences in rule ordering have less
effect on parsing time when parsing log files 22 \& 62--68, because of the
different distribution of log lines in those log files.  A careful
examination of \graphref{Parsing time of optimal and reverse orderings
relative to shuffled ordering graph} shows that, for the first log file
only, optimal and reverse orderings perform identically: this is because
the hits field of each rule is zero for the first log file, so optimal and
reverse orderings produce identical rule orderings.  For the first log
file, shuffled ordering is the most efficient of the three, but that is
accidental and cannot be relied upon.

@@

\showgraph{build/graph-file-size-optimal-shuffle-reverse}{Parsing time
plotted against log file size for optimal, shuffled, and reverse
orderings}{Parsing time plotted against log file size for optimal,
shuffled, and reverse orderings}

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important for this architecture because it is
anticipated that the typical parser will have a large ruleset.  The full
\parsername{} ruleset has \numberOFrules{} rules, whereas the minimum
ruleset required to parse the \numberOFlogFILES{} log files has
\numberOFrulesMINIMUM{} rules, \numberOFrulesMINIMUMpercentage{} of the
full ruleset.  The full ruleset is larger because \parsername{} is tested
with \numberOFlogFILESall{} log files; testing with more log files
increases the chance of finding bugs in the parser or new complications to
be overcome.  The \numberOFlogFILES{} log files were each parsed 10 times
using the minimum ruleset, and the mean parsing times compared to those
generated using the full ruleset: the percentage parsing time increase when
using the full ruleset instead of the minimal ruleset for optimal,
shuffled, and reverse orderings is shown in \graphref{Percentage parsing
time increase when using the maximum ruleset instead of the minimum
ruleset}, with mean and standard deviation in \tableref{Percentage parsing
time increase when using the maximum ruleset instead of the minimum ruleset
table}.  Note that in both \graphref{Percentage parsing time increase when
using the maximum ruleset instead of the minimum ruleset} and
\tableref{Percentage parsing time increase when using the maximum ruleset
instead of the minimum ruleset table}, for each ordering the parsing time
using the maximum ruleset is compared to the ordering's parsing time using
the minimum ruleset: the three orderings are not compared to a common
baseline.

@@

The optimal ordering shows a mean increase of just
\input{build/include-full-ruleset-vs-minimum-ruleset-mean} in parsing time
for a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
These results show that both the architecture and \parsername{} scale
extremely well as the ruleset increases in size, and that optimally
ordering the rules makes an important contribution to this scalability.

@@

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the effect that not caching compiled regexes has on parser
performance, with mean and standard deviation in \tableref{Increase in
parsing time when not caching compiled regexes table}.  For typical log
files, the mean increase in parsing time when not caching compiled regexes
is \input{build/include-cached-regexes-vs-discarded-regexes-mean}.  Caching
compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement: the framework
compiles each rule's regex when the ruleset is loaded, and uses the
compiled regex instead of the source regex when recognising log lines.  As
seen previously, log files 22 \& 62--68 do not suffer such a large increase
in parsing time when the optimisation is disabled; this is because, on
average, fewer rules are used so fewer regexes are compiled per log line
for those log files.  The increase in parsing time when parsing the first
log file is much greater than for the other log files; again, this is
because every rule's hits field starts at zero, so optimal ordering is less
efficient than usual, the mean number of rules used for each log line is
higher than usual, and more regexes will need to be compiled when
recognising each log line.

@@

The optimisations described in this chapter have addresses the process of
recognising log lines, but have not optimised actions at all.  Optimisation
efforts have concentrated on recognition of log lines for two reasons:

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still accounts for two thirds of the parser's
        execution time.  \Graphref{Percentage of parsing time spent
        recognising log lines graph} shows the percentage of parsing time
        spent recognising log lines for each of the \numberOFlogFILES{} log
        files, with mean and standard deviation shown in
        \tableref{Percentage of parsing time spent recognising log lines
        table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  To measure how
        long recognition of log lines takes, a full test run was performed
        as previously described in \sectionref{parser efficiency}, but with
        one difference: actions were not invoked when a log line was
        recognised, so the parsing time did not include the time normally
        taken by invocation of actions; the mean parsing times for each log
        file were subtracted from the mean parsing time from a normal test
        run to obtain the parsing time taken by actions.

        Optimal ordering reduces parsing time by
        \input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
        relative to shuffled ordering; actions occupy
        \input{build/include-percentage-time-spent-processing-log-lines-mean}
        of parsing time, less if any optimisations are disabled, so
        optimising actions could not possibly provide as big a performance
        increase as optimally ordering rules.  Similarly, caching compiled
        regexes provides an
        \input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}
        reduction in parsing time; not invoking actions at all reduces the
        most optimised parsing time by less than half that.  If processing
        of recognised log lines was optimised to 1\% of its original
        parsing time, it would be only slightly more effective than optimal
        ordering, but it would be vastly harder to implement; optimising
        actions would not reduce parsing time enough to justify the amount
        of effort required.
